---
output: html_document
---

Input: location(x,y), accuracy, timestamp
Output: place_id

#Analysis:

**Assumption:**
- Generally, place_id is most closely related to 'location+accuracy.'
- Since place_id is steady in space, and its open time is also decided. But the 'time' will be an auxiliary for 'accuracy.'

- Look at the sample submission, for each event id, there are 3 place_id, and according to the problem introduction, it's to return a ranking list rather than only 1 result? 'your task is to return a ranked list of the most likely places. '

###Sample submission
```{r}
sample <- fread("~/PycharmProjects/kaggle-project/facebook/sample_submission.csv")
head(sample)
```



###Load required packages
```{r, message = FALSE, warning = FALSE}
library(data.table) #reading in the data
library(dplyr) #dataframe manipulation
library(ggplot2) #viz
library(ranger) #the random forest implementation
library(plotly) #3D plotting
library(tidyr) #dataframe manipulation
library(FNN) #k nearest neighbors algorithm
library(xgboost)
```



###Load data, count total #, view columns

- Raw data has 29118021 rows, so it's better to firstly observe a sample.
- x and y are from 0 to 10, so the unit is km.
- accuracy has singular value, seems unit is 100 percent

```{r}
fb <- fread("~/PycharmProjects/kaggle-project/facebook/train.csv", integer64 = "character", showProgress = FALSE)
nrow(fb)
```

```{r}
head(fb, 3)
```

```{r}
summary(fb)
```

*How to represent accuracy?*


###Abstract sample area

**Assumption:**
- Since there are more than 100,000 places located in a 10 km by 10 km square.
  In this 0.25*0.25 area, there would supposed to be 625 unique place_id on average.


```{r}
fb %>% filter(x >1, x <1.25, y >2.5, y < 2.75) -> fb_s
nrow(fb_s)  
```

```{r}
fb %>% filter(x >3, x <3.25, y >2.5, y < 2.75) -> fb_s2
nrow(fb_s2)  
```

##Observe data
###place_id
**Since target is to classify place_id, 1st to observe place_id**


**Assumption:**
- From the figure 1&2, can see the trend of sorted place_id are almost same via different sample area.
- From figure 3, some events are in the same place_id class, but there are continusely increasing place ids between 2 id like transition bridge. That maybe a wrong classification.
```{r}
par(mfrow=c(3,1))
plot(sort(fb_s$place_id))
plot(sort(fb_s2$place_id))
plot(sort(fb_s2$place_id)[0:2000])
```

*As to ass-2, from figures: seem no relation between place_id and accuracy*

```{r}
par(mfrow=c(2,1))
r_pla_aur=sort(fb_s2$place_id,index.return=TRUE)
plot(r_pla_aur$x[0:2000])
d=r_pla_aur$ix[0:2000]
plot(fb_s2[d,"accuracy"])
```

```{r}
par(mfrow=c(2,1))
r_pla_aur=sort(fb_s2$place_id,index.return=TRUE)
plot(r_pla_aur$x[100:150])
d=r_pla_aur$ix[100:150]
plot(fb_s2[d,"accuracy"])
```

```{r}
par(mfrow=c(2,1))
r_pla_aur=sort(fb_s2$place_id,index.return=TRUE)
plot(r_pla_aur$x[500:750])
d=r_pla_aur$ix[500:750]
plot(fb_s2[d,"accuracy"])
```

*To count place_id*

**Assumption:**
- In this sample area, count by place_id, the 140 largest id seems to be correct class, but as estimation, there should be about 625 valid ids.

```{r}
nrow(fb_s %>% count(place_id))
```

```{r}
sort((fb_s %>% count(place_id))$n, decreasing = T)
```




###time
```{r}
fb_s$hour = (fb_s$time/60) %% 24
fb_s$weekday = (fb_s$time/(60*24)) %% 7
fb_s$month = (fb_s$time/(60*24*30)) %% 12 #month-ish
fb_s$year = fb_s$time/(60*24*365)
fb_s$day = (fb_s$time/(60*24)) %% 365
head(fb_s)
summary(fb_s)
```



##Train model
**Split data**

Since fb_s has 17710 rows, after sorted by time, split the data by 0.9 vs 0.1, so take the earlier 16000 events as training data, and the remaining to be valid data, and the 16000th data is time==713568, so we use 7.1e5 to be the filter.

```{r}
nrow(fb_s)
```

```{r}
sort(fb_s$time)[16000]
```

```{r}
small_train = fb_s[fb_s$time < 7.1e5,]
small_val = fb_s[fb_s$time >= 7.1e5,] 
```

**visualize 2D: small_train**

Visualize small train data by x,y, colored by place_id, since there are some overlap data, will use time or accuracy to sepreate. 

**Assumption:**
- There supposed to be unique (x,y) will have only 1 place_id, or to say, within the scope of a cluster, the (x,y)s share the same place id.

```{r, fig.height = 8, fig.width = 10}
ggplot(small_train, aes(x, y )) +
    geom_point(aes(color = place_id)) + 
    theme_minimal() +
    theme(legend.position = "none") +
    ggtitle("Check-ins colored by place_id")
```


**visualize 3D: small_train with place_id count>500**
**z=hour**

Count by place_id, will use the largest 8 group of place id to be colored.

```{r}
sort((small_train %>% count(place_id))$n, decreasing = T)[0:140]
```

**Observation:**
- From the 3D z=hour, place_id is determined by (x,y), will note change with time.
- There are overlap maybe by mistaken place_id, such as, in the green cluster, there are pink, orange and meat colors. These data should be modified or cleaned.
- Almost every cluster has dense and sparse area. This maybe another potential feature.
- Some clusters have singular points, such as gray and orange cluster. 
- Each place_is has its own open time, some are whole-day, some late night, etc.

**Assumption:**
- According to the problem introduction, 'Inconsistent and erroneous location data', for example, for the orange labeled point x located in the meat scope, the label is correct, but the location should not be there, but when the algorithm's leanring, this point will be identified as meat area. So is given a new data, it has the same location and accuracy, it should be classified as orange rather than meat. How?

```{r, fig.height = 8, fig.width = 8}
small_train %>% count(place_id) %>% filter(n > 500) -> ids
  #if n>200, warning: n too large, allowed maximum for palette Set2 is 8
small_trainz = small_train[small_train$place_id %in% ids$place_id,]

plot_ly(data = small_trainz, x = x , y = y, z = hour, color = place_id,  type = "scatter3d", mode = "markers", marker=list(size= 5)) %>% layout(title = "Place_id's by position and Time of Day")
```

- As to z=week, we can see different weekdays have different number of events, some days are busy while others are not.

**z=week**
```{r}
plot_ly(data = small_trainz, x = x , y = y, z = weekday, color = place_id,  type = "scatter3d", mode = "markers", marker=list(size= 5)) %>% layout(title = "Place_id's by position and Day of Week")

```


##Random Forest

**Count unique place_id**
```{r}
length(unique(small_train$place_id))
```

**Ignore fewer place_id**
```{r}
small_train %>% count(place_id) %>% filter(n > 3) -> ids
small_train = small_train[small_train$place_id %in% ids$place_id,]
```

#optimal weights for scaling your variables since knn is sensitive to the magnitutde of variables@¥: s,l,w
```{r}
summary(small_train)
```

```{r}
s = 2
l = 125
w = 500

create_matrix = function(train) {
    cbind(s*train$y,
          train$x,
          train$hour/l,
          train$weekday/w,
          train$year/w,
          train$month/w,
          train$time/(w*60*24*7))
    }

X = create_matrix(small_train)
X_val = create_matrix(small_val)

```

**KNN**
```{r}
model_knn = FNN::knn(train = X, test = X_val, cl = small_train$place_id, k = 15)

preds <- as.character(model_knn)
truth <- as.character(small_val$place_id)
mean(truth == preds)
```

```{r}
head(X)
```


**Random Forest**
```{r}
set.seed(131L)
small_train$place_id <- as.factor(small_train$place_id) # ranger needs factors for classification
model_rf <- ranger(place_id ~ x + y + accuracy + hour + weekday + month + year,
                   small_train,
                   num.trees = 100,
                   write.forest = TRUE,
                   importance = "impurity")


pred = predict(model_rf, small_val)
pred = pred$predictions
accuracy = mean(pred == small_val$place_id) 

accuracy
```

**Visualize RF accuracy**
It does seem that the correctly identified check-ins are more "clustered" while the wrongly identified ones are more uniformly distributed but other than that no clear patters here.
```{r}
small_val$Correct = (pred == small_val$place_id)

ggplot(small_val, aes(x, y )) +
    geom_point(aes(color = Correct)) + 
    theme_minimal() +
    scale_color_brewer(palette = "Set1")
```

**look at what kind of id's our random forest gets wrong**
We see below that our model is doing actually really great on the more popular id's(more blue on the right). However it loses when it looks at id's that appear only a few times. 
```{r, fig.width = 12}
#reordering the levels based on counts:
small_val$place_id <- factor(small_val$place_id,
                             levels = names(sort(table(small_val$place_id), decreasing = TRUE)))

small_val %>% 
    ggplot(aes(x = place_id)) + geom_bar(aes(fill = Correct)) + 
    theme_minimal() +
    theme(axis.text.x = element_blank()) +
    ggtitle("Prediction Accuracy by ID and Popularity") +
    scale_fill_brewer(palette = "Set1")
```

**importance of our variables**
1. `y` variable is more important than the `x`
This means that the `y` axis is a better predictior of `place_id` and the random forest figures this out on its own. 
2. `hour` and other time features are also good predictiors but less so than the spatial features - this makes sense since the location of a check-in should be more important than the time of the check-in.
3. Accuracy is a bit misterious since we don't get an actual definition for it, but at least the model tells us it's somewhat important.
```{r}
data.frame(as.list(model_rf$variable.importance)) %>% gather() %>% 
    ggplot(aes(x = reorder(key, value), y = value)) +
    geom_bar(stat = "identity", width = 0.6, fill = "grey") +
    coord_flip() +
    theme_minimal() +
    ggtitle("Variable Importance (Gini Index)") +
    theme(axis.title.y = element_blank()) 

```





